Postural developments mediate children's visual access to social information

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
naming events redo

```{r, libraries}
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=TRUE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(xtable)
library(lubridate)
library(langcog)
library(ggplot2)
library(ggthemes)
library(egg) # plot arranging
library(lme4)
library(gridExtra)
library(feather)
```
# Introduction

@@

To formalize these observations, we fit a generalized linear model to the proportion of faces infants saw in each posture and orientation (as detected by MTCNN), with participant's age, orientation, and posture as independent variables. A summary of the coefficients of a model with only main effects (and no interactions) model can be found in Table 3.  When we did include interaction terms between age, posture, and orientation, age no longer remained a significant predictor (b = `r round(faceGLMOut$coef[2,1],2)`, SE = `r round(faceGLMOut$coef[2,2],2)`, z = `r round(faceGLMOut$coef[2,3],2)` , p =`r round(faceGLMOut$coef[2,4],2)`).  We found the same pattern of results for the OpenPose face detections, omitted here for brevity. We also found the same pattern with respect to infant's visual access to hands (see Table 4); including interaction terms also reduced any main effect of age on the proportion of hands detected (b = `r round(wristGLMOut$coef[2,1],2)`, SE = `r round(wristGLMOut$coef[2,2],2)`, z = `r round(wristGLMOut$coef[2,3],2)` , p =`r round(wristGLMOut$coef[2,4],2)`). Thus, these results suggest that infants' access to social information is heavily influenced by their postural and locomotive development.

@@

aroundNamings_Orient_Wrists <- d %>%
  group_by(subid) %>%
  mutate(detections = wristOP) %>%
  do(summarize.naming(.)) %>%
  group_by(subid, orientation) %>%
  summarize(propWristsDetected = mean(detections)) %>%
  filter(!is.na(propWristsDetected)) %>%
  filter(!is.na(orientation)) %>%
  group_by(orientation) %>%
  multi_boot_standard(col = "propWristsDetected")

@@

p1=ggplot(aroundNamings_Posture_Faces, aes(x = posture, y = mean, col = posture)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ylab("Prop. faces") +
  xlab("") +
  ylim(0,.3) +
  theme_few() + 
  theme(legend.position="none")

p2=ggplot(aroundNamings_Orient_Faces, aes(x = orientation, y = mean, col = orientation)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ylab("Prop. faces") +
  xlab("") +
  ylim(0,.3) +
  theme_few() +
  theme(legend.position="none")

p3=ggplot(aroundNamings_Posture_Wrists, aes(x = posture, y = mean, col = posture)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ylab("Prop. wrists") +
  xlab("") +
  ylim(0,.3) +
  theme_few() +
  theme(legend.position="none")

p4=ggplot(aroundNamings_Orient_Wrists, aes(x = orientation, y = mean, col = orientation)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ylab("Prop. wrists") +
  xlab("") +
  ylim(0,.3) +
  theme_few() +
  theme(legend.position="none")

ggarrange(p1,p2, p3, p4)
```
## Access to Faces and Hands During Labeling Events
Finally, we analyzed how face and wrists detections changed during object labeling events as a function of infant's posture and orientation.  Specfically, we analyzed a four-second window around each labeling event (e.g., "Look at the [zem]!"); these labeling events were hand-annotated and automatically synchronized with the frame-by-frame face detections. We again found that infants' posture and orientations impacted the degree to which they saw their caregiver's face and wrist during a labeling event; infants who were sitting or standing were more likely to have access to this social information.

# General Discussion
We used a head-mounted camera to explore how children's postural and locomotive development directly impacts their access to social information, here operationalized as the presence of the faces and wrists of their caregiver. We found that children's posture and orientation towards their caregiver changed systematically across age, and that both of these factors dramatically impacted the proportion of faces and wrists that were avaliable in the child's visual field. Thus, infants' postural and locomotive developments are mediating factors that explain some of the age-related changes in the proportion of faces and hands avaliable in infants' visual field. Broadly, this work suggests that motoric developments mediate how infants experience their visual world and the social information in it: infants that are sitting and standing have a different view of their world, the people in it, and the actions that are being performed.

This work also deploys novel advancements in computer vision to the study of developmental psychology. The field of object detection and recognition has advanced dramatically in the past five years since the re-birth of deep learning algorithms [@krizhevsky2012imagenet], creating a new generation of algorithmic tools. These tools are substantially better equipped to deal with noisier, more complicated datasets and can extract richer and more detailed information. Videos from the infant perspective provided substantial challenges (e.g., partially occluded faces) for the classic models of face detection (e.g., ViolaJones) [@viola2004robust].  Further, as the headcam technologies employed here were inexpensive (~$60 a camera) and the computer vision algorithms freely avaliable, this method is a promising avenue for quanitfying the visual and social information avaliable to infant learners.

Thus, we suggest that the combined use of these new tools can be leveraged to understand the changing infant perspecive on the visual world and the implications of these changes for both linguistic, cognitive, and social development.

>>>>

```{r, libraries}
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=TRUE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(xtable)
library(lubridate)
library(langcog)
library(ggplot2)
library(ggthemes)
library(egg) # plot arranging
library(lme4)
library(gridExtra)
library(feather)
library(assertthat)
```
# Introduction

@@

To formalize these observations, we fit a generalized linear model to the proportion of faces infants saw in each posture and orientation (as detected by MTCNN), with participant's age, orientation, and posture as independent variables. A summary of the coefficients of a model with only main effects (and no interactions) model can be found in Table 3.  When we did include interaction terms between age, posture, and orientation, age no longer remained a significant predictor (b = `r round(faceGLMOut$coef[2,1],2)`, SE = `r round(faceGLMOut$coef[2,2],2)`, z = `r round(faceGLMOut$coef[2,3],2)` , p = `r round(faceGLMOut$coef[2,4],2)`).  We found the same pattern of results for the OpenPose face detections, omitted here for brevity. We also found the same pattern with respect to infant's visual access to hands (see Table 4); including interaction terms also eliminated any main effect of age on the proportion of hands detected (b = `r round(wristGLMOut$coef[2,1],2)`, SE = `r round(wristGLMOut$coef[2,2],2)`, z = `r round(wristGLMOut$coef[2,3],2)`, p = `r round(wristGLMOut$coef[2,4],2)`). Thus, these results suggest that infants' access to social information is heavily influenced by their postural and locomotive development.

@@

aroundNamings_Orient_Wrists <- d %>%
  group_by(subid) %>%
  mutate(detections = wristOP) %>%
  do(summarize.naming(.)) %>%
  group_by(subid, orientation) %>%
  summarize(propWristsDetected = mean(detections)) %>%
  filter(!is.na(propWristsDetected)) %>%
  filter(!is.na(orientation)) %>%
  group_by(orientation) %>%
  multi_boot_standard(col = "propWristsDetected") 
```

p1=ggplot(aroundNamings_Posture_Faces, aes(x = posture, y = mean, col = posture)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ylab("Prop. faces") +
  xlab("") +
  ylim(0,.25) +
  theme_few() + 
  theme(legend.position="none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p2=ggplot(aroundNamings_Orient_Faces, aes(x = orientation, y = mean, col = orientation)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ylab("Prop. faces") +
  xlab("") +
  ylim(0,.25) +
  theme_few() +
  theme(legend.position="none")  +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p3=ggplot(aroundNamings_Posture_Wrists, aes(x = posture, y = mean, col = posture)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ylab("Prop. wrists") +
  xlab("") +
  ylim(0,.25) +
  theme_few() +
  theme(legend.position="none")  +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p4=ggplot(aroundNamings_Orient_Wrists, aes(x = orientation, y = mean, col = orientation)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ylab("Prop. wrists") +
  xlab("") +
  ylim(0,.25) +
  theme_few() +
  theme(legend.position="none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggarrange(p1,p2, p3, p4)
```

```{r echo=FALSE, include=FALSE}
## Break down by before/after naming
### First for wrists
beforeNamings_Wrists <- d %>%
  group_by(subid) %>%
  mutate(detections = wristOP) %>%
  do(summarize.naming(.,window = c(-2,0))) %>%
  group_by(subid, age.grp)  %>%
  summarise(avgDetections = mean(detections, na.rm=TRUE))  %>%
  mutate(namingTime = "beforeNaming")

namings_Wrists <- d %>%
  group_by(subid) %>%
  mutate(detections = wristOP) %>%
  do(summarize.naming(.,window = c(0,2))) %>%
  group_by(subid, age.grp)  %>%
  summarise(avgDetections = mean(detections, na.rm=TRUE)) %>%
  mutate(namingTime = "afterNaming") %>%
  bind_rows(beforeNamings_Wrists)

### Now for faces
beforeNamings_Faces <- d %>%
  group_by(subid) %>%
  mutate(detections = faceMT) %>%
  do(summarize.naming(.,window = c(-2,0))) %>%
  group_by(subid, age.grp)  %>%
  summarise(avgDetections = mean(detections, na.rm=TRUE))  %>%
  mutate(namingTime = "beforeNaming")

namings_Faces <- d %>%
  group_by(subid) %>%
  mutate(detections = faceMT) %>%
  do(summarize.naming(.,window = c(0,2))) %>%
  group_by(subid, age.grp)  %>%
  summarise(avgDetections = mean(detections, na.rm=TRUE)) %>%
  mutate(namingTime = "afterNaming") %>%
  bind_rows(beforeNamings_Faces)

# Plot if you want to look
# p1=ggplot(namings_Faces, aes(x = factor(age.grp), y = avgDetections, col = age.grp)) +
#   geom_boxplot() +
#   ylab("Prop. faces") +
#   xlab("") +
#   theme_few() +
#   theme(legend.position="none") +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
#   facet_grid(~namingTime)
# 
# p2=ggplot(namings_Wrists, aes(x = factor(age.grp), y = avgDetections, col = age.grp)) +
#   geom_boxplot() +
#   ylab("Prop. wrists") +
#   xlab("") +
#   theme_few() +
#   theme(legend.position="none") +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
#   facet_grid(~namingTime)

# ggarrange(p1,p2) # not shown in .rmd

## output for text below \
naming_Wrists_Summary <-  namings_Wrists %>%
  spread(namingTime, avgDetections) %>%
  mutate(diffWrists = afterNaming - beforeNaming) %>%
  group_by(age.grp) %>%
  summarize(mean = mean(diffWrists), sd=sd(diffWrists))

naming_Faces_Summary <-  namings_Faces %>%
  spread(namingTime, avgDetections) %>%
  mutate(diffFaces = afterNaming - beforeNaming) %>%
  group_by(age.grp) %>%
  summarize(mean = mean(diffFaces), sd=sd(diffFaces)) # no room for SD maybe but still informative

@@

## Access to Faces and Hands During Labeling Events
Finally, we analyzed how face and wrist detections changed during object labeling events as a function of infant's posture and orientation. Specfically, we analyzed a four-second window around each labeling event (e.g., "Look at the [zem]!"); these labeling events were hand-annotated and automatically synchronized with the frame-by-frame face detections. We found that infants' posture and orientations impacted the degree to which they saw their caregiver's face and wrist during a labeling event; infants who were sitting or standing were more likely to have access to this social information. However, this did not manifest in a direct change the presence of wrists/hands before vs. after naming events (average difference in proportion of wrists detected, 8 m.o. = `r round(naming_Wrists_Summary$mean[1],3)`, 12 m.o. = `r round(naming_Wrists_Summary$mean[2],3)`, 16 m.o. = `r round(naming_Wrists_Summary$mean[3],3)`) or in proportion of faces detected  (average difference in proportion of faces detected, 8 m.o. = `r round(naming_Faces_Summary$mean[1],3)`, 12 m.o. = `r round(naming_Faces_Summary$mean[2],3)`, 16 m.o. = `r round(naming_Faces_Summary$mean[3],3)`).

# General Discussion
We used a head-mounted camera to explore how children's postural and locomotive development directly impacts their access to social information, here operationalized as the presence of the faces and wrists of their caregiver. We found that children's posture and orientation towards their caregiver changed systematically across age, and that both of these factors dramatically impacted the proportion of faces and wrists that were available in the child's visual field. Thus, infants' postural and locomotive developments are mediating factors that explain some of the age-related changes in the proportion of faces and wrists/hands that are visually available to infants. Broadly, this work suggests that motoric developments mediate how infants experience their visual world and the social information in it: infants that are sitting and standing have a different view of their world, the people in it, and the actions that are being performed.

This work also deploys novel advancements in computer vision to the study of developmental psychology. The field of object detection and recognition has advanced dramatically in the past five years since the re-birth of deep learning algorithms [@krizhevsky2012imagenet], creating a new generation of algorithmic tools. These tools are substantially better equipped to deal with noisier, more complicated datasets and can extract richer and more detailed information. Videos from the infant perspective provided substantial challenges (e.g., partially occluded faces) for the classic models of face detection (e.g., ViolaJones) [@viola2004robust].  Further, as the headcam technologies employed here were inexpensive and the computer vision algorithms freely available, this method is a promising avenue for quantifying the visual and social information available to infant learners.

Broadly, we suggest that the combined use of these new tools can be leveraged to understand the changing infant perspective on the visual world and the implications of these changes for linguistic, cognitive, and social development.
